---
title: "Data Science Homework 3"
author: "Laura Lynch"
date: "10/14/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(viridis)
library(dplyr)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Problem 1
```{r}
library(p8105.datasets)
data("instacart")

instacart
```
The instacart dataset reflects the business of the instacart online groceryordering service. Order data includes the ailse and deartment information for the product as well as exactly when the order was placed and how frequently the customer uses the service. There are `r nrow(distinct(instacart, department))` departments and `r nrow(distinct(instacart, product_id))` items to choose from which have been ordered by `r nrow(distinct(instacart, user_id))` different customers.

 * There are `r nrow(distinct(instacart, aisle_id))` aisles and aisles most ordered from are:
```{r}
instacart %>%
  select(aisle, product_id) %>%
  group_by(aisle) %>% 
  summarize(n_aisle = n()) %>%
  arrange(desc(n_aisle)) %>%
  head(3) %>%
  knitr::kable()
```

 
 * This plot reflects the number of items ordered in each aisle:
 
```{r}
instacart %>%
  group_by(aisle) %>% 
  summarize(n_aisle = n()) %>%
  filter(n_aisle > 10000) %>%
  arrange(desc(n_aisle)) %>%
  ggplot(aes(x = aisle, y = n_aisle, color = aisle)) +
  geom_point() +
  labs(
    title = "Instacart Plot",
    x = "aisle",
    y = "number of orders") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```

 * This table shows the three most popular items in each of the aisles
```{r}
instacart %>%
  filter(aisle %in% c("packaged vegetables fruits", "dog food care", "baking ingredients")) %>%
  group_by(aisle) %>% 
  count(product_name, name = "product_count") %>%
  mutate(count_ranking = min_rank(desc(product_count))) %>%
  filter(min_rank(count_ranking) < 4) %>%
knitr::kable()
```


 * This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:

```{r}
instacart %>%
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_time = mean(order_hour_of_day)) %>%
  mutate(order_dow = recode(order_dow, "0" = "Monday", "1" = "Tuesday", "2" = "Wednesday", "3" = "Thursday", "4" = "Friday", "5" = "Saturday", "6" = "Sunday")) %>%
  pivot_wider(
   names_from = "order_dow", 
   values_from = "mean_time") %>%
  knitr::kable(digits = 3)
```



Problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 
```

Clean the `BRFSS Dataset`:
```{r}
tidy_brfss = 
brfss_smart2010 %>%
  janitor::clean_names() %>%   
  rename(state = locationabbr, county = locationdesc) %>% 
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good","Very good", "Excellent"))) %>%
  separate(county, into = c("state2", "county"), sep = 3) %>%
  mutate(county = stringr::str_replace(county, "- ", "")) %>%
  select(-state2) 

tidy_brfss
```


 * In 2002, which states were observed at 7 or more locations? 
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
  filter(year == 2002) %>%
  group_by(state) %>%
   summarize(
    n_counties = n_distinct(county)) %>%
  filter(n_counties > 6)
```
Answer: CT, FL, MA< NC, NJ, PA
 
 * What about in 2010?
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
  filter(year == 2010) %>%
  group_by(state) %>%
   summarize(
    n_counties = n_distinct(county)) %>%
  filter(n_counties > 6)
```
Answer: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA

 * Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state, make a spagetti plot to visualize the change:
```{r}
tidy_brfss %>% 
  filter(response == "Excellent") %>%
    group_by(year, state, county) %>%
    summarize(mean_data_value = mean(data_value, na.rm = TRUE)) %>%
    select(year, state, county, mean_data_value) %>%
  ggplot(aes(x = year, y = mean_data_value, color = state)) + geom_line() + 
  labs(
    title = "Mean Value over Time by State",
    x = "Year",
    y = "Mean Value")
```

 *Two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State:
```{r}
response_2006 = 
  tidy_brfss %>%
        filter(year == 2006, state == "NY") %>%
  ggplot(aes(x = response, y = data_value, color = response)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "2006 Responses",
    x = "Response",
    y = "Value")
  
response_2010 = 
  tidy_brfss %>%
        filter(year == 2010, state == "NY") %>%
  ggplot(aes(x = response, y = data_value, color = response)) + 
  geom_point(alpha = .5)  + 
  labs(
    title = "2010 Responses",
    x = "Response",
    y = "Value")

(response_2006 + response_2010)

```

Problem 3

Import data:
```{r}
accel_data = read_csv(file = "./data/accel_data.csv")

accel_data
```

Clean and tidy:
```{r}
accel_data_tidy =
  accel_data %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_min",
    values_to = "activity_count") %>%
    mutate(activity_min = stringr::str_replace(activity_min, "activity_", "")) %>%
    mutate(activity_min = as.numeric(activity_min)) %>%
    mutate(weekday = day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),
           weekend = day %in% c("Saturday", "Sunday")) 

accel_data_tidy
```
Describe the resulting dataset:
This dataset is tracking a study particpants activity across the `r nrow(distinct(accel_data_tidy, day_id))` days of the study. The columns track the week of the study, the day of the week as well as if it is a weekday or a weekend. The participants activity is level is trarcked every minute. Over the course of the study his average acticity count was `r mean(pull(accel_data_tidy, activity_count))`, median acticity was `r median(pull(accel_data_tidy, activity_count))` and standard deviation was `r sd(pull(accel_data_tidy, activity_count))`.


2) Traditional analyses of accelerometer data focus on the total activity over the day. 
Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals: 
```{r}
accel_data_tidy %>%
  group_by(day_id) %>%
    summarize(daily_activity = sum(activity_count)) %>%
  knitr::kable()
```

```{r}
# visualize data from table in #2:
accel_data_tidy %>%
   group_by(week, day_id) %>%
    summarize(daily_activity = sum(activity_count)) %>%
  ggplot(aes(x = day_id, y = daily_activity)) + geom_polygon() + scale_x_continuous(
    breaks = c(7, 14, 21, 29), 
    labels = c("Week 1", "Week 2", "Week 3", "Week 4"))  + labs(
     title = "Summary of Activity by Day",
      x = "Day",
      y = "Activity Level")
```

* Apparent trends: From the start of the study until the end of the study, the participants general activity level declined. Compared to the first half of the study, during the second half of the study, the participant had longer periods of activity that were below the mean. During the first half of the study, the participant had regularly spaced, activity exertions that seemed to fall off mid-way through.


Plot of activity throughout the day:
```{r}
accel_data_tidy %>%
  group_by(day_id) %>%
    ggplot(aes(x = activity_min, y = activity_count, group = day_id, color= day)) +
      geom_smooth(alpha = .5)  + 
        scale_x_continuous(
        breaks = c(60, 120, 180, 240, 300, 360, 420, 480, 540, 600, 660, 720, 780, 840, 900, 960, 1020, 1080, 1140, 1200, 1260, 1320, 1380, 1440), 
labels = c("0100h","0200h", "0300h", "0400h", "0500h", "0600h", "0700h", "0800h", "0900h", "1000h", "1100h", "1200h", "1300h", "1400h", "1500h", "1600h", "1700h", "1800h", "1900h", "2000h", "2100h", "2200h", "2300h", "0000h"))+
  labs(
    title = "Activity throughout the day",
    x = "Hour of the day",
    y = "Activity Level"
  ) +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```
# Patterns or conclusions: It looks like the participant generally starts their day around 6:00. The activity level begins to increase then and plateaus around 8:00 am where it remains constant until around 6:00 pm, perhaps this is when the participant is at work. On Monday, Friday and Saturday there is another exertion and then steady decline in activity, the other days the decline begins from the plateu.There does appear to be one Saturday with no activity which could be a device malfunction or an extremely restful day for the participant.