---
title: "Data Science Homework 3"
author: "Laura Lynch"
date: "10/06/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(viridis)
library(dplyr)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Problem 1
```{r}
library(p8105.datasets)
data("instacart")

instacart %>% view
```
The instacart dataset reflects the business of the instacart online groceryordering service. Order data includes the ailse and deartment information for the product as well as exactly when the order was placed and how frequently the customer uses the service. There are `r nrow(distinct(instacart, department))` departments and `r nrow(distinct(instacart, product_id))` items to choose from which have been ordered by `r nrow(distinct(instacart, user_id))` different customers.

 * There are `r nrow(distinct(instacart, aisle_id))` aisles and aisles most ordered from are:
```{r}
instacart %>%
  select(aisle, product_id) %>%
  group_by(aisle) %>% 
  summarize(n_aisle = n()) %>%
  arrange(desc(n_aisle)) %>%
  head(3) %>%
  knitr::kable()
```

 
 * This plot reflects the number of items ordered in each aisle
 
```{r}
instacart %>%
  group_by(aisle) %>% 
  summarize(n_aisle = n()) %>%
  filter(n_aisle > 10000) %>%
  arrange(desc(n_aisle)) %>%
  ggplot(aes(x = aisle, y = n_aisle, color = aisle)) +
  geom_point() +
  labs(
    title = "Instacart Plot",
    x = "aisle",
    y = "number of orders") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```

 * This table shows the three most popular items in each of the aisles
```{r}
instacart %>%
  filter(aisle %in% c("packaged vegetables fruits", "dog food care", "baking ingredients")) %>%
  group_by(aisle) %>% 
  count(product_name, name = "product_count") %>%
  mutate(count_ranking = min_rank(desc(product_count))) %>%
  filter(min_rank(count_ranking) < 4) %>%
knitr::kable()
```


 * This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:

```{r}
instacart %>%
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_time = mean(order_hour_of_day)) %>%
  pivot_wider(
   names_from = "order_dow", 
   values_from = "mean_time") %>%
  knitr::kable() 
# round mean
# numbers to days of week
```
# mutate(order_dow = day.name[order_dow]) %>%


Problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 %>% view
```

Clean the `BRFSS Dataset`:
```{r}
brfss_smart2010 %>%
  janitor::clean_names() %>%   
  rename(state = locationabbr, county = locationdesc) %>% #melt county?
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  knitr::kable() 
```

 * In 2002, which states were observed at 7 or more locations? 
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
  filter(year == 2002) %>%
  group_by(state) %>%
   summarize(
    n_counties = n_distinct(county)) %>%
  filter(n_counties > 6)
```
 
 * What about in 2010?
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
  filter(year == 2010) %>%
  group_by(state) %>%
   summarize(
    n_counties = n_distinct(county)) %>%
  filter(n_counties > 6)
```

 * Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
  filter(response == "Excellent") %>%
  group_by(year, state, county) %>%
  summarize(mean_data_value = mean(data_value)) %>%
  select(year, state, county, mean_data_value)

```

 * Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
    group_by(year, state, county) %>%
    summarize(mean_data_value = mean(data_value)) %>%
ggplot(aes(x = year, y = mean_data_value)) + 
  geom_line(aes(color = state), alpha = .5) + 
  labs(
    title = "Mean Value over Time by State",
    x = "Year",
    y = "Mean Value"
  )

# what does value mean?
```


 *Two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State
```{r}
response_2006 = 
  brfss_smart2010 %>% 
    janitor::clean_names() %>%
      rename(state = locationabbr, county = locationdesc) %>%
        filter(year == 2006, state == "NY") %>%
          filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
          mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  ggplot(aes(x = data_value, y = response, color = response)) + 
  geom_point(alpha = .5) 
  
response_2010 = 
  brfss_smart2010 %>% 
    janitor::clean_names() %>%
      rename(state = locationabbr, county = locationdesc) %>%
        filter(year == 2010, state == "NY") %>%
          filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
          mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  ggplot(aes(x = data_value, y = response, color = response)) + 
  geom_point(alpha = .5)  

(response_2006 + response_2010)
```


Problem 3

Import data:
```{r}
accel_data = read_csv(file = "./data/accel_data.csv")
  view(accel_data)
```


Clean and tidy:
```{r}
accel_data  %>% 
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_min",
    values_to = "activity_count") %>%
    mutate(activity_min = stringr::str_replace(activity_min, "activity_", "")) %>%
    mutate(activity_min = as.numeric(activity_min)) %>%
    mutate(weekday = day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),
           weekend = day %in% c("Saturday", "Sunday")) 
```
# Describe the resulting dataset (e.g. what variables exist, how many observations, etc).


2) Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. 
```{r}

```
# Are any trends apparent?


3)Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 
```{r}

```
# Describe in words any patterns or conclusions you can make based on this graph
