---
title: "Data Science Homework 3"
author: "Laura Lynch"
date: "10/06/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(viridis)
library(dplyr)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Problem 1
```{r}
library(p8105.datasets)
data("instacart")

instacart
```
The instacart dataset reflects the business of the instacart online groceryordering service. Order data includes the ailse and deartment information for the product as well as exactly when the order was placed and how frequently the customer uses the service. There are `r nrow(distinct(instacart, department))` departments and `r nrow(distinct(instacart, product_id))` items to choose from which have been ordered by `r nrow(distinct(instacart, user_id))` different customers.

 * There are `r nrow(distinct(instacart, aisle_id))` aisles and aisles most ordered from are:
```{r}
instacart %>%
  select(aisle, product_id) %>%
  group_by(aisle) %>% 
  summarize(n_aisle = n()) %>%
  arrange(desc(n_aisle)) %>%
  head(3) %>%
  knitr::kable()
```

 
 * This plot reflects the number of items ordered in each aisle:
 
```{r}
instacart %>%
  group_by(aisle) %>% 
  summarize(n_aisle = n()) %>%
  filter(n_aisle > 10000) %>%
  arrange(desc(n_aisle)) %>%
  ggplot(aes(x = aisle, y = n_aisle, color = aisle)) +
  geom_point() +
  labs(
    title = "Instacart Plot",
    x = "aisle",
    y = "number of orders") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```

 * This table shows the three most popular items in each of the aisles
```{r}
instacart %>%
  filter(aisle %in% c("packaged vegetables fruits", "dog food care", "baking ingredients")) %>%
  group_by(aisle) %>% 
  count(product_name, name = "product_count") %>%
  mutate(count_ranking = min_rank(desc(product_count))) %>%
  filter(min_rank(count_ranking) < 4) %>%
knitr::kable()
```


 * This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:

```{r}
instacart %>%
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_time = mean(order_hour_of_day)) %>%
  pivot_wider(
   names_from = "order_dow", 
   values_from = "mean_time") %>%
  knitr::kable(digits = 3)

# mutate(order_dow = recode(order_dow, 0 = "Monday", 2 = "Tuesday", 3 = "Wednesday", 4 = "Thursday", 5= "Friday", 6 = "Saturday")
```
# mutate(order_dow = day.name[order_dow]) %>%


Problem 2

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 
```

Clean the `BRFSS Dataset`:
```{r}
tidy_brfss = 
brfss_smart2010 %>%
  janitor::clean_names() %>%   
  rename(state = locationabbr, county = locationdesc) %>% 
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good","Very good", "Excellent"))) %>%
  separate(county, into = c("state2", "county"), sep = 3) %>%
  mutate(county = stringr::str_replace(county, "- ", "")) %>%
  select(-state2) 
```
# mutate(county = stringr::str_replace(county, " - ", "")) %>%
# mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good","Very good", "Excellent"))) %>%
# mutate(response = factor(response, levels = c("Poor", "Fair","Good","Very good", "Excellent")))

 * In 2002, which states were observed at 7 or more locations? 
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
  filter(year == 2002) %>%
  group_by(state) %>%
   summarize(
    n_counties = n_distinct(county)) %>%
  filter(n_counties > 6)
```
 
 * What about in 2010?
```{r}
brfss_smart2010 %>%
    janitor::clean_names() %>%
    rename(state = locationabbr, county = locationdesc) %>%
  filter(year == 2010) %>%
  group_by(state) %>%
   summarize(
    n_counties = n_distinct(county)) %>%
  filter(n_counties > 6)
```

 * Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
```{r}
tidy_brfss %>% 
  filter(response == "Excellent") %>%
    group_by(year, state, county) %>%
    summarize(mean_data_value = mean(data_value, na.rm = TRUE)) %>%
    select(year, state, county, mean_data_value) %>%
  ggplot(aes(x = year, y = mean_data_value, color = state)) + geom_line() + 
  labs(
    title = "Mean Value over Time by State",
    x = "Year",
    y = "Mean Value")
```

 *Two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State
```{r}
response_2006 = 
  tidy_brfss %>%
        filter(year == 2006, state == "NY") %>%
  ggplot(aes(x = response, y = data_value, color = response)) + 
  geom_point(alpha = .5) 
  
response_2010 = 
  tidy_brfss %>%
        filter(year == 2010, state == "NY") %>%
  ggplot(aes(x = response, y = data_value, color = response)) + 
  geom_point(alpha = .5)  

(response_2006 + response_2010)

# + facet_grid(day_of_tx ~ outcome)
```
         
    redundant:      filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
          mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%

          filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% 
          mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%


Problem 3

Import data:
```{r}
accel_data = read_csv(file = "./data/accel_data.csv")

accel_data
```

Clean and tidy:
```{r}
accel_data_tidy =
  accel_data %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_min",
    values_to = "activity_count") %>%
    mutate(activity_min = stringr::str_replace(activity_min, "activity_", "")) %>%
    mutate(activity_min = as.numeric(activity_min)) %>%
    mutate(weekday = day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),
           weekend = day %in% c("Saturday", "Sunday")) 

accel_data_tidy
```
# Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
This dataset is tracking a study particpants activity across the `r nrow(distinct(accel_data_tidy, day_id))` days of the study. The columns track the week of the study, the day of the week as well as if it is a weekday or a weekend. The participants activity is level is trarcked every minute. Over the course of the study his average acticity count was `r mean(pull(accel_data_tidy, activity_count))`, median acticity was `r median(pull(accel_data_tidy, activity_count))` and standard deviation was `r sd(pull(accel_data_tidy, activity_count))`.


2) Traditional analyses of accelerometer data focus on the total activity over the day. 
Using your tidied dataset, aggregate accross minutes to create a 
total activity variable for each day, 
and create a table showing these totals. 
```{r}
accel_data_tidy %>%
  group_by(day_id) %>%
    summarize(daily_activity = mean(activity_count)) %>%
  knitr::kable()
```
```{r}
# Any trends?
accel_data_tidy %>%
  group_by(week, day_id) %>%
    summarize(daily_activity = mean(activity_count)) %>%
  ggplot(aes(x = day_id, y = daily_activity, color = week)) + geom_polygon() + scale_x_continuous(
    breaks = c(7, 14, 21, 29), 
    labels = c("Week 1", "Week 2", "Week 3", "Week 4"))
```


From the starrt of the study until the end of the study, the participants generalactivity level declined


Plot of activity throughout the day:
```{r}
accel_data_tidy %>%
  group_by(day_id) %>%
    ggplot(aes(x = activity_min, y = activity_count, group = day_id, color= day)) +
      geom_smooth(alpha = .5)  + 
        scale_x_continuous(
        breaks = c(60, 120, 180, 240, 300, 360, 420, 480, 540, 600, 660, 720, 780, 840, 900, 960, 1020, 1080, 1140, 1200, 1260, 1320, 1380, 1440), 
labels = c("0100h","0200h", "0300h", "0400h", "0500h", "0600h", "0700h", "0800h", "0900h", "1000h", "1100h", "1200h", "1300h", "1400h", "1500h", "1600h", "1700h", "1800h", "1900h", "2000h", "2100h", "2200h", "2300h", "0000h"))+
  labs(
    title = "Activity throughout the day",
    x = "Hour of the day",
    y = "Activity Level"
  ) +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```
# Describe in words any patterns or conclusions you can make based on this graph

